---
title: "The Organoid Project"
output: html_notebook
author: Tim Sit
---

Here is the analysis done in `R` for The Organoid Project in 2018.

# Spike timing tiling coefficient 

Here I implement the tiling coefficient introduced by Cutts and Eglen (2014) as a way to quantify the correlation between pairs of spike trains.

I have translated their code from C to Matlab, to test if it worked properly, I shall first test it on a hypothetical pair of spike trains: 
spike train 1 (spike times)

[1, 3, 5, 7]

spike train 2 (spike times)

[2, 3, 6, 10]

Where the total recording time is 10s (From 1 - 11)

Using the spike timing tiling coefficient from the `sjemea` package, from here: https://github.com/sje30/sjemea

```{r}
library(sjemea)
a <- c(236.454000000000,
237.216480000000,
288.159480000000,
295.043560000000,
356.737440000000,
379.664360000000)

b <- c(236.454000000000,
236.694080000000,
237.216520000000,
288.159480000000,
295.003560000000,
295.043560000000,
328.342160000000,
356.737440000000,
379.664320000000,
691.200600000000)

tiling_coefficient <- tiling.corr(a, b, dt = 0.01, rec.time = c(1/25000, 720))
```


Also note that some people have implemented a slightly modified version of this in `meaRtools`

```{r}
library(meaRtools)

a = c(1, 0, 1, 0, 1, 0, 1, 0, 0, 0)
b = c(0, 1, 1, 0, 0, 1, 0, 0, 0, 1)
sttc(a, b, dt = 1, rec_time = 10)

```

